---
title: "AI Tool Use, Academic Impact, and Work Readiness Among College Students"
author: "Zainab Safdary"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 7.5, fig.height = 4.5)

# ---- Check packages BEFORE loading (gives one clear message if any are missing)
needed  <- c("readr", "dplyr", "tidyr", "ggplot2", "knitr", "sandwich", "lmtest", "MASS")
missing <- needed[!vapply(needed, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing) > 0) {
  stop("Run this in the Console, then restart R (Cmd+Shift+0) and knit again:\n",
       "install.packages(c(", paste0('"', missing, '"', collapse = ", "), "))")
}

library(MASS)      # loaded before dplyr so dplyr::select is not masked
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(sandwich)
library(lmtest)

theme_set(
  theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5, color = "grey35"),
          axis.title = element_text(face = "bold"),
          panel.grid.minor = element_blank())
)
pal <- c(AIS = "#2C7FB8", WRS = "#41AB5D")
```

```{r helpers, include=FALSE}
# ---- Small helper functions used throughout the report ----------------------

stars <- function(p) ifelse(p < .01, "***", ifelse(p < .05, "**", ifelse(p < .10, "*", "")))

fmt_p <- function(p) ifelse(p < .001, "< .001", sprintf("%.3f", p))

# For use inside sentences: "p < .001" or "p = 0.023"
p_text <- function(p) ifelse(p < .001, "p < .001", sprintf("p = %.3f", p))

# Readable names for regression terms (":" is replaced first on purpose)
pretty_term <- function(x) {
  x <- gsub("\\(Intercept\\)", "Intercept", x)
  x <- gsub(":", " × ", x)
  x <- gsub("AI_Usage_Freq", "AI usage frequency", x)
  x <- gsub("Usage_Level", "Usage level = ", x)
  x <- gsub("Course_Group", "Course = ", x)
  x <- gsub("College_Year", "Year = ", x)
  x <- gsub("English_PrimaryNo", "English not primary language", x)
  x
}

# Coefficients with HC3 heteroskedasticity-robust standard errors
robust_coefs <- function(model) {
  ct <- coeftest(model, vcov. = vcovHC(model, type = "HC3"))
  data.frame(term = rownames(ct), est = ct[, 1], se = ct[, 2], p = ct[, 4],
             row.names = NULL)
}

# Side-by-side regression table: "estimate*** (robust SE)" in each cell
reg_table <- function(models, caption) {
  cols <- lapply(names(models), function(nm) {
    rc  <- robust_coefs(models[[nm]])
    out <- data.frame(term = rc$term,
                      value = sprintf("%.3f%s (%.3f)", rc$est, stars(rc$p), rc$se))
    names(out)[2] <- nm
    out
  })
  tab <- Reduce(function(a, b) full_join(a, b, by = "term"), cols)
  tab$term <- pretty_term(tab$term)

  fit <- data.frame(
    term = c("Observations", "R²", "Adjusted R²"),
    vapply(models, function(m) c(as.character(nobs(m)),
                                 sprintf("%.3f", summary(m)$r.squared),
                                 sprintf("%.3f", summary(m)$adj.r.squared)),
           character(3)),
    check.names = FALSE
  )
  tab <- rbind(tab, fit)
  tab[is.na(tab)] <- ""
  names(tab)[1] <- ""
  kable(tab, caption = caption, align = c("l", rep("c", length(models))))
}

# Cronbach's alpha, written out so no extra package is needed
cronbach_alpha <- function(items) {
  items <- as.matrix(items)
  k <- ncol(items)
  (k / (k - 1)) * (1 - sum(apply(items, 2, var)) / var(rowSums(items)))
}
```

```{r load-data}
# Looks next to this .Rmd file first, then in Desktop/Senior Project
paths     <- c("survey_data.csv", "~/Desktop/Senior Project/survey_data.csv")
data_path <- paths[file.exists(paths)][1]
if (is.na(data_path)) {
  stop("survey_data.csv not found. Save this .Rmd in the same folder as the CSV.")
}

raw <- read_csv(data_path, show_col_types = FALSE)
```

```{r clean}
survey <- raw %>%
  mutate(
    # Fix a typo in the age responses
    Age_Range = recode(Age_Range, "Obove 18" = "Above 18"),
    Under_18  = Age_Range == "Under 18",

    # "Languages and Linguistics" has only 4 students, so it is grouped with "other"
    Course_Group = case_when(
      Course_Type %in% c("other...", "Languages and Linguistics") ~ "Other",
      TRUE ~ Course_Type),
    Course_Group = factor(Course_Group,
                          levels = c("Mathematics + Computer Science", "Social Sciences",
                                     "Writing and Composition", "Other")),

    College_Year    = factor(College_Year,
                             levels = c("Freshman", "Sophomore", "Junior", "Senior")),
    English_Primary = factor(Primary_English, levels = c("Yes", "No")),

    # Usage as categories (for testing whether the effect is linear)
    Usage_Level = factor(AI_Usage_Freq, levels = 0:4,
                         labels = c("0 times", "1-2", "3-4", "5-6", "More than 6")),

    # Rebuild the composites from their items to confirm the file is correct
    AIS_check = 0.4 * Learning_vs_Completion + 0.3 * Grade_Improvement_Pct +
                0.3 * Extra_Time_Activities,
    WRS_check = 0.4 * AI_Career_Importance + 0.3 * Productivity_Loss_Pct +
                0.3 * Work_Quality_Comparison,

    # Alternative versions used in the robustness section
    AIS_equal    = (Learning_vs_Completion + Grade_Improvement_Pct + Extra_Time_Activities) / 3,
    WRS_equal    = (AI_Career_Importance + Productivity_Loss_Pct + Work_Quality_Comparison) / 3,
    WRS_reversed = 0.4 * AI_Career_Importance + 0.3 * (1 - Productivity_Loss_Pct) +
                   0.3 * Work_Quality_Comparison
  )

stopifnot(max(abs(survey$AIS - survey$AIS_check)) < 0.001,
          max(abs(survey$WRS - survey$WRS_check)) < 0.001)

n_total <- nrow(survey)
n_minor <- sum(survey$Under_18)
```

# Introduction

This study tests three hypotheses:

- **H1.** Students who use AI tools more often report a higher Academic Impact Score (AIS).
- **H2.** Students who use AI tools more often report a higher Work Readiness Score (WRS).
- **H3.** The relationship between AI use and these outcomes differs by course type.

# Data and Measures

The data come from an online survey of `r n_total` college students collected in
October 2024. All participants gave consent. Survey responses were converted to
numeric scales before analysis:

- **AI usage frequency** is coded 0–4 (0 times, 1–2, 3–4, 5–6, more than 6).
- **Single items** are rescaled to run from 0 (lowest response) to 1 (highest).
- **Academic Impact Score (AIS)** = 0.4 × learning vs. completion + 0.3 × grade
  improvement + 0.3 × uses saved time differently.
- **Work Readiness Score (WRS)** = 0.4 × AI career importance + 0.3 × productivity
  loss without AI + 0.3 × work quality with AI.
- Four additional attitude items are measured on 1–5 scales.

"Languages and Linguistics" (4 students) is combined with "Other" because a group
that small cannot be estimated reliably.

## Sample characteristics

```{r demographics}
demo_vars <- c(College_Year = "College year", Course_Group = "Course type",
               Age_Range = "Age", Primary_English = "English is primary language")

demo <- bind_rows(lapply(names(demo_vars), function(v) {
  survey %>%
    count(Category = as.character(.data[[v]])) %>%
    mutate(Variable = demo_vars[[v]],
           Percent = sprintf("%.1f%%", 100 * n / sum(n)))
})) %>%
  dplyr::select(Variable, Category, N = n, Percent)

demo$Variable[duplicated(demo$Variable)] <- ""
kable(demo, caption = "Sample characteristics", align = c("l", "l", "r", "r"))
```

```{r descriptives}
num_vars <- c(
  AI_Usage_Freq           = "AI usage frequency (0–4)",
  AIS                     = "Academic Impact Score (0–1)",
  WRS                     = "Work Readiness Score (0–1)",
  Learning_vs_Completion  = "  Learning vs. completion",
  Grade_Improvement_Pct   = "  Grade improvement",
  Extra_Time_Activities   = "  Uses saved time differently (0/1)",
  AI_Career_Importance    = "  AI career importance",
  Productivity_Loss_Pct   = "  Productivity loss without AI",
  Work_Quality_Comparison = "  Work quality with AI",
  AI_Learning_Alignment   = "AI learning alignment (1–5)",
  AI_Usage_Likelihood     = "AI usage likelihood (1–5)",
  AI_Good_Student         = "AI and being a good student (1–5)",
  AI_Thinking_Change      = "AI changes thinking (1–5)"
)

survey %>%
  dplyr::select(all_of(names(num_vars))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "value") %>%
  group_by(var) %>%
  summarise(N = sum(!is.na(value)), Mean = mean(value, na.rm = TRUE),
            SD = sd(value, na.rm = TRUE), Min = min(value, na.rm = TRUE),
            Median = median(value, na.rm = TRUE), Max = max(value, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(Variable = num_vars[var]) %>%
  arrange(match(var, names(num_vars))) %>%
  dplyr::select(Variable, N, Mean, SD, Min, Median, Max) %>%
  kable(digits = 2, caption = "Table 2. Descriptive statistics")
```

```{r usage-plot}
ggplot(survey, aes(x = Usage_Level)) +
  geom_bar(fill = pal[["AIS"]], alpha = 0.85, width = 0.7) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  labs(title = "Students AI Tools Usage Frequency",
       x = "AI usage frequency", y = "Number of students")
```

```{r score-hist}
survey %>%
  dplyr::select(AIS, WRS) %>%
  pivot_longer(everything(), names_to = "Score", values_to = "Value") %>%
  ggplot(aes(x = Value, fill = Score)) +
  geom_histogram(binwidth = 0.1, boundary = 0, color = "white", alpha = 0.9) +
  facet_wrap(~ Score, labeller = as_labeller(c(AIS = "Academic Impact Score",
                                               WRS = "Work Readiness Score"))) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(title = "Distribution of Composite Scores",
       x = "Score (0–1)", y = "Number of students")
```

```{r miss-plot, fig.height=3.5}
survey %>%
  count(Miss_AI_Most) %>%
  mutate(Miss_AI_Most = reorder(Miss_AI_Most, n)) %>%
  ggplot(aes(x = n, y = Miss_AI_Most)) +
  geom_col(fill = "grey45", width = 0.65) +
  geom_text(aes(label = n), hjust = -0.3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "What Students Would Miss Most Without AI",
       x = "Number of students", y = NULL)
```

# Measurement Quality

Before using the composite scores, I check whether their items measure a common
idea (internal consistency) and whether the chosen weights are reasonable.

```{r reliability}
ais_items <- survey %>%
  dplyr::select(Learning_vs_Completion, Grade_Improvement_Pct, Extra_Time_Activities)
wrs_items <- survey %>%
  dplyr::select(AI_Career_Importance, Productivity_Loss_Pct, Work_Quality_Comparison)

alpha_ais <- cronbach_alpha(ais_items)
alpha_wrs <- cronbach_alpha(wrs_items)

# First principal component: how strongly each item reflects the shared factor
pca_weights <- function(items) {
  pc   <- prcomp(items, scale. = TRUE)
  load <- pc$rotation[, 1]
  if (sum(load) < 0) load <- -load
  list(loadings = load,
       weights  = load / sum(load),
       var_expl = summary(pc)$importance[2, 1])
}
pca_ais <- pca_weights(ais_items)
pca_wrs <- pca_weights(wrs_items)

kable(data.frame(
  Score = c("AIS", "WRS"),
  `Cronbach's alpha` = round(c(alpha_ais, alpha_wrs), 3),
  `Variance explained by 1st component` = sprintf("%.1f%%", 100 * c(pca_ais$var_expl, pca_wrs$var_expl)),
  check.names = FALSE
), caption = "Table 3. Internal consistency of composite scores")

kable(data.frame(
  Score = rep(c("AIS", "WRS"), each = 3),
  Item = c(names(ais_items), names(wrs_items)),
  `Chosen weight` = c(0.4, 0.3, 0.3, 0.4, 0.3, 0.3),
  `PCA-implied weight` = round(c(pca_ais$weights, pca_wrs$weights), 2),
  check.names = FALSE
), caption = "Table 4. Chosen weights compared with data-driven (PCA) weights")
```

Cronbach's alpha is `r round(alpha_ais, 2)` for AIS and `r round(alpha_wrs, 2)`
for WRS. Both are below the conventional 0.70 benchmark. This means the items are
related but capture somewhat different experiences, so the composites should be
read as *indices* that summarize several outcomes, not as measures of one single
trait. With only three items per score, lower alpha values are expected. Section 6
checks whether the main results depend on how the scores are weighted.

# Results

## Correlations

```{r corr-heatmap, fig.height=5.5, fig.width=7.5}
corr_vars <- c(AI_Usage_Freq = "AI usage", AIS = "AIS", WRS = "WRS",
               AI_Learning_Alignment = "Learning alignment",
               AI_Usage_Likelihood = "Usage likelihood",
               AI_Good_Student = "Good student",
               AI_Thinking_Change = "Thinking change")

cm <- cor(survey[, names(corr_vars)], method = "spearman")
dimnames(cm) <- list(corr_vars, corr_vars)

as.data.frame(as.table(cm)) %>%
  ggplot(aes(Var1, Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", Freq)), size = 3.5) +
  scale_fill_gradient2(low = "#D73027", mid = "white", high = "#2C7FB8",
                       limits = c(-1, 1), name = "Spearman ρ") +
  labs(title = "Correlations Among Key Variables", x = NULL, y = NULL) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

ct_ais <- cor.test(survey$AI_Usage_Freq, survey$AIS, method = "spearman", exact = FALSE)
ct_wrs <- cor.test(survey$AI_Usage_Freq, survey$WRS, method = "spearman", exact = FALSE)
```

Spearman correlations are used because AI usage is an ordered category. AI usage
is strongly correlated with both AIS (ρ = `r round(ct_ais$estimate, 2)`,
`r p_text(ct_ais$p.value)`) and WRS (ρ = `r round(ct_wrs$estimate, 2)`,
`r p_text(ct_wrs$p.value)`).

## Scores by level of AI use

```{r means-by-usage}
usage_means <- survey %>%
  pivot_longer(c(AIS, WRS), names_to = "Score", values_to = "Value") %>%
  group_by(Score, Usage_Level) %>%
  summarise(n = n(), mean = mean(Value), se = sd(Value) / sqrt(n()), .groups = "drop") %>%
  mutate(lo = mean - qt(0.975, n - 1) * se,
         hi = mean + qt(0.975, n - 1) * se)

ggplot(usage_means, aes(x = Usage_Level, y = mean, color = Score, group = Score)) +
  geom_line(linewidth = 1, position = position_dodge(width = 0.3)) +
  geom_pointrange(aes(ymin = lo, ymax = hi), position = position_dodge(width = 0.3)) +
  scale_color_manual(values = pal) +
  labs(title = "Average Scores by AI Usage Frequency",
       subtitle = "Points are group means; bars are 95% confidence intervals",
       x = "AI usage frequency", y = "Mean score (0–1)", color = NULL)
```

The "5-6" group has only `r sum(survey$AI_Usage_Freq == 3)` students, which is why
its confidence interval is wide.

## Regression models

Four models are estimated for each outcome:

1. **Simple:** AI usage only.
2. **Controlled:** adds course type, college year, and whether English is the
   student's primary language.
3. **Categorical:** treats usage as separate categories to test whether each step
   up has the same effect.
4. **Interaction:** lets the effect of AI usage differ by course type (H3).

All standard errors are HC3 heteroskedasticity-robust.
\* p < .10, \*\* p < .05, \*\*\* p < .01.

```{r models}
controls <- "Course_Group + College_Year + English_Primary"

fit_set <- function(outcome) {
  list(
    Simple      = lm(as.formula(paste(outcome, "~ AI_Usage_Freq")), data = survey),
    Controlled  = lm(as.formula(paste(outcome, "~ AI_Usage_Freq +", controls)), data = survey),
    Categorical = lm(as.formula(paste(outcome, "~ Usage_Level +", controls)), data = survey),
    Interaction = lm(as.formula(paste(outcome, "~ AI_Usage_Freq * Course_Group +",
                                      "College_Year + English_Primary")), data = survey)
  )
}
ais_models <- fit_set("AIS")
wrs_models <- fit_set("WRS")
```

```{r ais-table}
reg_table(ais_models, "Table 5. Predictors of Academic Impact Score (robust SEs)")
```

```{r wrs-table}
reg_table(wrs_models, "Table 6. Predictors of Work Readiness Score (robust SEs)")
```

```{r model-tests}
nested_p <- function(small, big) anova(small, big)$`Pr(>F)`[2]

usage_effect <- function(models, outcome) {
  full    <- models$Controlled
  reduced <- lm(as.formula(paste(outcome, "~", controls)), data = survey)
  r2_full <- summary(full)$r.squared
  r2_red  <- summary(reduced)$r.squared
  std_fit <- lm(as.formula(paste0("scale(", outcome, ") ~ scale(AI_Usage_Freq) +", controls)),
                data = survey)
  rc <- robust_coefs(full)
  rc <- rc[rc$term == "AI_Usage_Freq", ]
  data.frame(
    Outcome = outcome,
    `b (per usage step)` = round(rc$est, 3),
    `95% CI` = sprintf("[%.3f, %.3f]", rc$est - 1.96 * rc$se, rc$est + 1.96 * rc$se),
    `Standardized β` = round(coef(std_fit)[["scale(AI_Usage_Freq)"]], 2),
    `Cohen's f²` = round((r2_full - r2_red) / (1 - r2_full), 2),
    `p: linear vs. categorical` = fmt_p(nested_p(full, models$Categorical)),
    `p: interaction with course` = fmt_p(nested_p(full, models$Interaction)),
    check.names = FALSE
  )
}

effects <- bind_rows(usage_effect(ais_models, "AIS"), usage_effect(wrs_models, "WRS"))
int_p_ais <- nested_p(ais_models$Controlled, ais_models$Interaction)
int_p_wrs <- nested_p(wrs_models$Controlled, wrs_models$Interaction)
kable(effects, caption = "Table 7. Size of the AI usage effect and model comparison tests")
```

**Reading Table 7.** The coefficient *b* is the change in score (on the 0–1 scale)
for each step up in usage frequency, holding the controls constant. The standardized β
shows the same effect in standard-deviation units. For Cohen's f², 0.02 is
conventionally small, 0.15 medium, and 0.35 large. A non-significant "linear vs.
categorical" test means treating usage as a single number is adequate. A
significant "interaction" test would mean the effect of AI depends on course type.

After controls, each step up in AI use is associated with a
`r round(effects$"b (per usage step)"[1], 3)` higher AIS and a
`r round(effects$"b (per usage step)"[2], 3)` higher WRS. The test for different effects by course type gives
`r p_text(int_p_ais)` for AIS and `r p_text(int_p_wrs)` for WRS, so the effect of AI
use `r ifelse(int_p_ais < .05, "does", "does not")` differ significantly by course
type for AIS and `r ifelse(int_p_wrs < .05, "does", "does not")` for WRS.

```{r coef-plot, fig.height=4.5}
coef_df <- bind_rows(
  robust_coefs(ais_models$Controlled) %>% mutate(Score = "AIS"),
  robust_coefs(wrs_models$Controlled) %>% mutate(Score = "WRS")
) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = pretty_term(term),
         lo = est - 1.96 * se, hi = est + 1.96 * se)

ggplot(coef_df, aes(x = est, y = term, color = Score)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = pal) +
  labs(title = "Controlled Model Estimates",
       subtitle = "Points are estimates; lines are 95% robust confidence intervals",
       x = "Effect on score (0–1 scale)", y = NULL, color = NULL)
```

```{r interaction-plot}
survey %>%
  pivot_longer(c(AIS, WRS), names_to = "Score", values_to = "Value") %>%
  ggplot(aes(x = AI_Usage_Freq, y = Value, color = Course_Group)) +
  geom_jitter(width = 0.12, height = 0, alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  facet_wrap(~ Score) +
  labs(title = "Does the AI Effect Differ by Course Type?",
       x = "AI usage frequency (0–4)", y = "Score (0–1)", color = "Course type") +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2))
```

## Attitude items: ordered logistic regression

The four 1–5 attitude items are ordered categories, so they are modeled with
proportional-odds (ordered logit) models instead of OLS. Each model includes the
same controls.

```{r ordinal}
ordinal_outcomes <- c(AI_Learning_Alignment = "AI learning alignment",
                      AI_Usage_Likelihood   = "AI usage likelihood",
                      AI_Good_Student       = "AI and being a good student",
                      AI_Thinking_Change    = "AI changes thinking")

ordinal_results <- bind_rows(lapply(names(ordinal_outcomes), function(y) {
  d   <- survey %>% mutate(.y = factor(.data[[y]], ordered = TRUE))
  fit <- polr(as.formula(paste(".y ~ AI_Usage_Freq +", controls)), data = d, Hess = TRUE)
  co  <- coef(summary(fit))["AI_Usage_Freq", ]
  p   <- 2 * pnorm(-abs(co[["t value"]]))
  data.frame(Outcome = ordinal_outcomes[[y]],
             `Odds ratio` = round(exp(co[["Value"]]), 2),
             `95% CI` = sprintf("[%.2f, %.2f]",
                                exp(co[["Value"]] - 1.96 * co[["Std. Error"]]),
                                exp(co[["Value"]] + 1.96 * co[["Std. Error"]])),
             p = fmt_p(p),
             check.names = FALSE)
}))

kable(ordinal_results,
      caption = "Table 8. Effect of one step up in AI usage on attitude items (ordered logit)")
```

An odds ratio above 1 means students who use AI more often are more likely to give
a higher response on that item. For example, an odds ratio of 2 means each step
up in usage doubles the odds of choosing a higher category.

## Differences across course types

```{r course-plot, fig.height=5}
survey %>%
  pivot_longer(c(AIS, WRS), names_to = "Score", values_to = "Value") %>%
  ggplot(aes(x = Course_Group, y = Value, fill = Score)) +
  geom_boxplot(alpha = 0.55, outlier.shape = NA) +
  geom_jitter(width = 0.15, height = 0, alpha = 0.4, size = 1.3) +
  facet_wrap(~ Score) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(title = "Scores by Course Type", x = NULL, y = "Score (0–1)") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r course-tests}
course_test <- function(outcome) {
  a  <- aov(as.formula(paste(outcome, "~ Course_Group")), data = survey)
  s  <- summary(a)[[1]]
  kw <- kruskal.test(as.formula(paste(outcome, "~ Course_Group")), data = survey)
  data.frame(Outcome = outcome,
             `ANOVA F` = round(s$`F value`[1], 2),
             `ANOVA p` = fmt_p(s$`Pr(>F)`[1]),
             `η²` = round(s$`Sum Sq`[1] / sum(s$`Sum Sq`), 3),
             `Kruskal-Wallis χ²` = round(kw$statistic[[1]], 2),
             `Kruskal-Wallis p` = fmt_p(kw$p.value),
             check.names = FALSE)
}
course_results <- bind_rows(course_test("AIS"), course_test("WRS"))
kable(course_results, caption = "Table 9. Do scores differ by course type?")

tukey <- TukeyHSD(aov(AIS ~ Course_Group, data = survey))$Course_Group
kable(round(tukey, 3), caption = "Table 10. Tukey pairwise comparisons for AIS")
```

The Kruskal-Wallis test does not assume normally distributed scores, so it serves
as a check on the ANOVA. The Kruskal-Wallis test gives
`r p_text(kruskal.test(AIS ~ Course_Group, data = survey)$p.value)` for AIS and
`r p_text(kruskal.test(WRS ~ Course_Group, data = survey)$p.value)` for WRS.

# Robustness Checks

## Sensitivity to score construction and sample

The controlled model is re-estimated with (a) equal weights instead of 0.4/0.3/0.3,
(b) productivity loss reverse-coded, which treats dependence on AI as *lower*
readiness, and (c) the `r n_minor` respondents under 18 excluded.

```{r sensitivity}
sens_spec <- list(
  "AIS, original weights"          = list(y = "AIS",          d = survey),
  "AIS, equal weights"             = list(y = "AIS_equal",    d = survey),
  "AIS, excluding under 18"        = list(y = "AIS",          d = filter(survey, !Under_18)),
  "WRS, original weights"          = list(y = "WRS",          d = survey),
  "WRS, equal weights"             = list(y = "WRS_equal",    d = survey),
  "WRS, productivity loss reversed"= list(y = "WRS_reversed", d = survey),
  "WRS, excluding under 18"        = list(y = "WRS",          d = filter(survey, !Under_18))
)

sens <- bind_rows(lapply(names(sens_spec), function(nm) {
  s  <- sens_spec[[nm]]
  m  <- lm(as.formula(paste(s$y, "~ AI_Usage_Freq +", controls)), data = s$d)
  rc <- robust_coefs(m)
  rc <- rc[rc$term == "AI_Usage_Freq", ]
  data.frame(Specification = nm,
             `AI usage b` = sprintf("%.3f%s", rc$est, stars(rc$p)),
             `Robust SE` = round(rc$se, 3),
             p = fmt_p(rc$p),
             N = nobs(m),
             `R²` = round(summary(m)$r.squared, 3),
             check.names = FALSE)
}))
kable(sens, caption = "Table 11. AI usage coefficient across alternative specifications")
```

*Interpretation: if the coefficient keeps the same sign and significance in every
row, the findings do not depend on the weighting choices or on including minors.
Comment especially on the reversed-WRS row: it shows how much the WRS result
depends on treating "I'd lose productivity without AI" as a sign of readiness.*

## Regression diagnostics

```{r diagnostics, fig.height=7, fig.width=8}
op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
plot(ais_models$Controlled, sub.caption = "AIS controlled model")
plot(wrs_models$Controlled, sub.caption = "WRS controlled model")
par(op)
```

```{r diag-tests}
diag_row <- function(m, name) {
  data.frame(Model = name,
             `Breusch-Pagan p (constant variance)` = fmt_p(bptest(m)$p.value),
             `Shapiro-Wilk p (normal residuals)` = fmt_p(shapiro.test(residuals(m))$p.value),
             `Max Cook's distance` = round(max(cooks.distance(m)), 3),
             check.names = FALSE)
}
kable(bind_rows(diag_row(ais_models$Controlled, "AIS"),
                diag_row(wrs_models$Controlled, "WRS")),
      caption = "Table 12. Assumption tests (p < .05 suggests a violation)")
```


